### 2024-03-11 paper/paper.qmd
added paper template and simulate section.

---
title: "Investigating the correlation between a vote for Biden or Trump based on a persons education, gender, race and age."
author: 
  - Fares Alkorani
  - Owais Zahid
thanks: "Code and data are available at: https://github.com/FFFiend/linear_model_investigation."
date: today
date-format: long
abstract: "In this paper, we discuss the correlation between various statistics of the respondents to the CES2020 election survey, and whether they voted for Trump or Biden. The statistics are: age, education, gender and race."
format: pdf
number-sections: true
bibliography: references.bib
---
# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Simulation
We first start with simulating the dataset, and recall that we are concerned with each voter's education level, gender, race and age. Our chosen sample size is 8000, and the categories for education and race have been named akin to the categories in the dataset. Thus, the header of the table obtained for the simulation is shown below:
```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false
# setting seed, prime number :P
set.seed(71)
library(dplyr)
# sample size 8000
num_obs <- 8000

# simulating the actual data
# using 8 categories for races as in the dataset, and ages 0-100.
us_political_preferences <- tibble(
  education = sample(0:4, size = num_obs, replace = TRUE),
  gender = sample(0:1, size = num_obs, replace = TRUE),
  race = sample(1:8, size = num_obs, replace = TRUE),
  age = sample(0:100, size = num_obs, replace=TRUE),
  support_prob = ((education + gender + race + age) / 7),
) |>
  mutate(
    supports_biden = if_else(runif(n = num_obs) < support_prob, "yes", "no"),
    education = case_when(
      education == 0 ~ "< High school",
      education == 1 ~ "High school",
      education == 2 ~ "Some college",
      education == 3 ~ "College",
      education == 4 ~ "Post-grad"
    ),
    gender = if_else(gender == 0, "Male", "Female"),
    race = case_when(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Hispanic",
      race == 4 ~ "Asian",
      race == 5 ~ "Native American",
      race == 6 ~ "Middle Eastern",
      race == 7 ~ "Two or more races",
      race == 8 ~ "Other"
    )
  ) |>
  select(-support_prob, supports_biden, gender, education, race, age)

# preview the simulated data
cat("Table 1. Header for the political preferences dataset.")
head(us_political_preferences)
```

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

## Diagnostics

\newpage

# References

### 2024-03-11 paper/paper.qmd
data section for paper

@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
-
-The remainder of this paper is structured as follows. @sec-data....
+We are concerned with assessing whether we can predict what candidate a new survey respondent is going to vote for, given their age, education, gender and race.
+Accordingly, we shall be simulating a similar scenario for the 2020 US presidential election below.
 
@@
 set.seed(71)
+library(knitr)
+library(tibble)
 library(dplyr)
@@
 cat("Table 1. Header for the political preferences dataset.")
-head(us_political_preferences)
+kable(head(us_political_preferences))
 ```
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+We will be using the 2020 Cooperative Election Study (CES) (CITATION TODO) as our dataset, which can be previewed below:
 
-Talk more about it.
+```{r}
+#| include: true
+#| warning: false
+#| message: false
+#| echo: false
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+library(dplyr)
+library(tidyverse)
 
-Talk way more about it.
+raw_data = read.csv("~/linear_model_investigation/data/ces2020_raw.parquet")
+kable(head(raw_data))
+```
+To elaborate further on the dataset starting with the ```votereg``` column, values 1 and 2 signifies whether a person has voted or not, respectively.
+Naturally, for entries with a value of 2 under this column, the corresponding ```CC20_410``` value is NA, and in the other case the values 1 and 2 signify a vote for Biden or Trump respectively.
+The naming of this column corresponds to the question in the survey from which the data for it was collected, which asks "For whom did you vote for President of the United States?" (citation needed), with a total of 6 options, where Biden and Trump correspond to the first two.
 
+Next, there are two values for gender: 1 and 2 corresponding to Male and Female.
+There are also a total of 8 categories under race corresponding to numbers 1 through 8, as well as 6 education categories ranging from "No HS" to "Post-grad".
+
+Since we are interested with ages and not respondent birth years, we may process the dataset as such to reflect this:
+
+```{r}
+#| include: true
+#| warning: false
+#| message: false
+#| echo: false
+
+raw_data$birthyr <- as.numeric(raw_data$birthyr)
+current_year <- as.numeric(year(as.Date(Sys.time())))
+
+# Subtract current year from birth year column to get age
+raw_data <- raw_data %>% mutate(birthyr = current_year - birthyr)
+# rename to age
+names(raw_data)[names(raw_data) == "birthyr"] <- "age"
+cat("Table 3. Header for 2020 US Election Survey data with the birthyr column converted to ages")
+kable(head(raw_data))
+
+# save data back into the csv.
+write.csv(raw_data,"~/linear_model_investigation/data/ces2020.parquet")
+```
+We shall now filter for non NA values within the ```CC20_410``` column, as we are only concerned with respondents that cast a vote, which also means we filter for entries with a ```votereg``` value of 1.
+Additionally, we shall map each education level and race value to its corresponding value from the survey, and we finally obtain a table as follows:
+
+```{r}
+#| include: true
+#| warning: false
+#| message: false
+#| echo: false
+# load in libraries.
+library(tidyverse)
+library(dplyr)
+library(rstanarm)
+library(modelsummary)
+
+# read and convert values in each column to integer.
+ces2020 <-
+  read_csv(
+    "~/linear_model_investigation/data/ces2020.parquet",
+    col_types =
+      cols(
+        "votereg" = col_integer(),
+        "CC20_410" = col_integer(),
+        "gender" = col_integer(),
+        "educ" = col_integer(),
+        "race" = col_integer(),
+        "age" = col_integer()
+      )
+  )
+
+# formatting data, mapping race ids to races.
+ces2020 <-
+  ces2020 |>
+  filter(votereg == 1,
+         CC20_410 %in% c(1, 2)) |>
+  mutate(
+    voted_for = if_else(CC20_410 == 1, "Biden", "Trump"),
+    voted_for = as_factor(voted_for),
+    gender = if_else(gender == 1, "Male", "Female"),
+    education = case_when(
+      educ == 1 ~ "No HS",
+      educ == 2 ~ "High school graduate",
+      educ == 3 ~ "Some college",
+      educ == 4 ~ "2-year",
+      educ == 5 ~ "4-year",
+      educ == 6 ~ "Post-grad"
+    ),
+    education = factor(
+      education,
+      levels = c(
+        "No HS",
+        "High school graduate",
+        "Some college",
+        "2-year",
+        "4-year",
+        "Post-grad"
+      )
+    ),
+    race = case_when(
+      race == 1 ~ "White",
+      race == 2 ~ "Black",
+      race == 3 ~ "Hispanic",
+      race == 4 ~ "Asian",
+      race == 5 ~ "Native American",
+      race == 6 ~ "Middle Eastern",
+      race == 7 ~ "Two or more races",
+      race == 8 ~ "Other"
+    ),
+    race = factor(
+      race,
+      levels = c(
+        "White",
+        "Black",
+        "Hispanic",
+        "Asian",
+        "Native American",
+        "Middle Eastern",
+        "Two or more races",
+        "Other"
+      )
+    )
+  ) |> select(voted_for, gender, education, race, age)
+cat("Table 4. Header for Processed 2020 US Election Survey data.")
+# preview data after mutation.
+kable(head(ces2020))
+```
 # Model
@@
 Our results are summarized in @tbl-modelresults.
+
+```{r}
+#| include: true
+#| warning: false
+#| message: false
+#| echo: false
+cat("results")
+political_preferences <-
+  readRDS(file = "~/linear_model_investigation/data/political_preferences.rds")
+
+# model summary
+final_table = modelsummary(
+  list(
+    "Support Biden" = political_preferences
+  ),
+  statistic = "mad"
+)
+kable(final_table)
+```
 

### 2024-03-11 scripts/01-download_data.R

@@ -20,5 +20,5 @@ ces2020 <-
   ) |>
   select(votereg, CC20_410, gender, educ, race, birthyr)
 
-write_csv(ces2020, "~/linear_model_investigation/data/ces2020.parquet")
+write_csv(ces2020, "~/linear_model_investigation/data/ces2020_raw.parquet")
 

### 2024-03-11 paper/paper.qmd
added formal model definition

@@
 ```
+TODO: complete bar chart section from the worked example.
 # Model
+Since we wish to predict how likely it is for a respondent to cast a vote for Biden based on their gender, age, education and race, we shall use the rstanarm (citation needed) library to
 
@@
 ## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+We define $y_i$ as the candidate the $ith$ respondent likely voted for, mapping 1 and 0 to Biden and Trump respectively.
+The variables $gender_i$, $age_i$, $race_i$ and $education_i$ are the selected attributes (with the same names, of) of the $ith$ respondent.
 
 \begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i|\pi_i, &\sim \mbox{Bern}(\pi_i) \\
+logit(\pi_i) = \beta_0 + \beta_1 \times gender_i + \beta_2 \times age_i + \beta_3 \times education_i + \beta_4 \times race_i \\ 
+\beta_0 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_4 &\sim \mbox{Normal}(0, 2.5)
 \end{align}
@@
 ### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+I actually have no idea what to expect lol
 

### 2024-03-16 scripts/02-clean_data.R
fixed ages

@@ -10,7 +10,7 @@
 library(dplyr)
 library(tidyverse)
 
-raw_data = read.csv("~/linear_model_investigation/data/ces2020.parquet")
+raw_data = read.csv("~/linear_model_investigation/data/ces2020_raw.parquet")
 raw_data
 
 raw_data$birthyr <- as.numeric(raw_data$birthyr)

### 2024-03-16 scripts/02-clean_data.R
fixed ages 2

@@ -14,8 +14,7 @@ raw_data = read.csv("~/linear_model_investigation/data/ces2020_raw.parquet")
 raw_data
 
 raw_data$birthyr <- as.numeric(raw_data$birthyr)
-current_year <- as.numeric(year(as.Date(Sys.time())))
-
+current_year <- 2020
 # Subtract current year from birth year column to get age
 raw_data <- raw_data %>% mutate(birthyr = current_year - birthyr)
 # rename to age

### 2024-03-16 scripts/03-test_data.R

@@ -32,3 +32,4 @@ sum(is.na(data$educ)) == 0
 
 # Test 7, check for NA values inside the gender column
 sum(is.na(data$gender)) == 0
+

### 2024-03-16 scripts/02-clean_data.R
Fixed scripts
Modified the relative paths in the scripts such that they can run

@@ -9,8 +9,9 @@
 # here we are converting the birth year column to age
 library(dplyr)
 library(tidyverse)
+library(arrow)
 
-raw_data = read.csv("~/linear_model_investigation/data/ces2020_raw.parquet")
+raw_data = read_parquet("data/ces2020_raw.parquet")
 raw_data
 
 raw_data$birthyr <- as.numeric(raw_data$birthyr)
@@ -19,8 +20,9 @@ current_year <- 2020
 raw_data <- raw_data %>% mutate(birthyr = current_year - birthyr)
 # rename to age
 names(raw_data)[names(raw_data) == "birthyr"] <- "age"
-head(raw_data)
+
+
 
 # save data back into the csv.
-write.csv(raw_data,"~/linear_model_investigation/data/ces2020.parquet")
+write_parquet(raw_data,"data/ces2020.parquet")
 

### 2024-03-16 scripts/01-download_data.R
test

@@ -20,5 +20,5 @@ ces2020 <-
   ) |>
   select(votereg, CC20_410, gender, educ, race, birthyr)
 
-write_csv(ces2020, "~/linear_model_investigation/data/ces2020_raw.parquet")
+write_csv(ces2020, "data/ces2020_raw.parquet")
 

### 2024-03-16 scripts/01-download_data.R
test

@@ -20,5 +20,5 @@ ces2020 <-
   ) |>
   select(votereg, CC20_410, gender, educ, race, birthyr)
 
-write_csv(ces2020, "data/ces2020_raw.parquet")
+write_csv(ces2020, "~/linear_model_investigation/data/ces2020_raw.parquet")
 

### 2024-03-16 paper/paper.qmd
fixed ages
fixed

@@
 raw_data$birthyr <- as.numeric(raw_data$birthyr)
-current_year <- as.numeric(year(as.Date(Sys.time())))
+current_year <- 2020
 
@@
 TODO: complete bar chart section from the worked example.
+
 # Model
@@
 #| message: false
+#| eval: true
 #| echo: false
@@
 # model summary
-final_table = modelsummary(
+modelsummary::modelsummary(
   list(
@@
 )
-kable(final_table)
 ```

### 2024-03-16 scripts/01-download_data.R

@@ -11,7 +11,6 @@
 # and filtering for the chosen explanatory variables and the response variable.
 library(dataverse)
 library(tidyverse)
-library(arrow)
 
 ces2020 <-
   get_dataframe_by_name(
@@ -22,5 +21,5 @@ ces2020 <-
   ) |>
   select(votereg, CC20_410, gender, educ, race, birthyr)
 
-write_parquet(ces2020, "data/ces2020_raw.parquet")
+write_csv(ces2020, "data/ces2020_raw.parquet")
 

### 2024-03-16 scripts/02-clean_data.R

@@ -9,9 +9,9 @@
 # here we are converting the birth year column to age
 library(dplyr)
 library(tidyverse)
-library(arrow)
 
-raw_data = read_parquet("data/ces2020_raw.parquet")
+
+raw_data = read_csv("data/ces2020_raw.parquet")
 raw_data
 
 raw_data$birthyr <- as.numeric(raw_data$birthyr)
@@ -24,5 +24,5 @@ names(raw_data)[names(raw_data) == "birthyr"] <- "age"
 
 
 # save data back into the csv.
-write_parquet(raw_data,"data/ces2020.parquet")
+write_csv(raw_data,"data/ces2020.parquet")
 

### 2024-03-17 paper/paper_fares.qmd
Updated PDF

---
title: "Deciphering Key Determinants of Biden Voter Support: Insights from Logistic Regression Analysis in the 2020 US Presidential Election"
author: 
  - Fares Alkorani
  - Owais Zahid
thanks: "Code and data are available at: https://github.com/FFFiend/linear_model_investigation."
date: today
date-format: long
abstract: "In this paper, we discuss the correlation between various statistics of the respondents to the CES2020 election survey, and whether they voted for Trump or Biden. The statistics are: age, education, gender and race."
format: pdf
number-sections: true
bibliography: references.bib
---
# Introduction

On November 4, 2020, the 59th United States presidential election took place between Joe Biden and Donald Trump, where Joe Biden was elected as the 46th president of the United States. Historically, there tend to be common demographics among voters who support the Democratic or Republican party and by extension, the presidential candidates from these parties. In addition, voter demographics are strategically considered in presidential campaigns to appeal to certain citizens.

As a result, in this paper the estimand we hope to explore is the likelihood that an American citizen will vote for Joe Biden in the 2020 presidental election given their income, gender, race, and age. We will do this by using the 2020 Cooperative Election Study (CES) [CITATION] which is a national online survey that was administered to a representative sample of American citizens pre and post-election, gathering demographic information from 61,000 respondents. Using 8000 responses from this survey, we construct a logistic regression model using income, gender, race, and age as predictor variables to see which demographics aligned with support for Joe Biden. We find that (**I NEED YOUR MOST RELEVANT FINDINGS HERE**). This information is important as it can allow us to understand the campaign strategies in the 2024 United States presidential election toward citizens with specific demographics, especially as Donald Trump and Joe Biden will likely be representatives of their respective parties for the upcoming presidential election.

The remainder of this paper is structured as follows: Section 2 discusses the data from the 2020 CES, Section 3 discusses the logistic regression model that was constructed from the 2020 CES data, Section 4 presents the results, and finally Section 5 discusses our findings and some weaknesses. (**MOSTLY FROM https://tellingstorieswithdata.com/04-writing_research.html#introduction-1, DUNNO IF IT NEEDS REPHRASING OR GIVEN ADDITIONAL DETAILS LATER**).

# Data {#sec-data}

## Data Source

In this paper, we will be using the 2020 Cooperative Election Study (CES) [CITATION], administered by YouGov, is a yearly online survey that gathers information regarding American adults and their political attitudes. This study was chosen because its dataset is publicly accessible, includes vote validation, and contains a diverse selection of attributes which make it possible to study how demographics of American adults correlates with preference for presidential candidate.

## Data Measurement

Most, if not all variables in the 2020 CES dataset correspond to a question asked in the CCES online questionnaire. Most respondents to this questionnaire are YouGov panelists, but other respondents were recruited from online surveys or other survey providers. [CITATION]

However, in order to gain a representative sample of all American adults, not all respondents of the CCES questionnaire end up in the final dataset [CITATION]. The representative sample was created by using matching, "is a methodology for selection of “representative” samples from non-randomly selected pools of respondent" [CITATION] which is appropriate in this case as the pool of respondents is primarily made of YouGov panelists - which is nonrandom.

Lastly, "individual records were matched to the Catalist database of registered voters in the United States" to validate the votes. However, only respondents with a high confidence of being matched in the Catalist database were included. Thus, some respondents may have been omitted from the final dataset in this regard.

## Data Features

```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false

library(dplyr)
library(tidyverse)

raw_data = read.csv("data/ces2020_raw.parquet")
kable(head(raw_data))
```

We use the R programming language [@citeR] and dataverse [CITATION] to access the CES dataset which initially begins with 61,000 respondents. From this dataset, we begin with the following variables as shown in the [FIRST TABLE CROSS REFERENCE]. The values from these variables from the questionnaire corresponding to each variable can be found in the Appendix along with the value it was numerically coded to in the original dataset [CROSS REFERENCE: TABLE 1].

* votereg: This is a binary variable where the values 1 and 2 signify whether a person is registered to vote or not, respectively.
* cc20_410: This variable measures who the respondent voted for as President of the United States if they voted (with the values 1 and 2 corresponding to a vote for Biden and Trump, respectively) and values 4 through 7 that distinguish non-voters
* gender: This is a binary variable which measures if the respondent is male (1) or female (2)
* educ: This is a nominal variable which measures the level of education the respondent has ranging from 1 - "Did not graduate from high school" to 6 - "Postgraduate degree"
* race: This is a nominal variable which numerically codes races from the questionnaire to numbers in the dataset
* birthyr: This variable measures the year of birth of a respondent.

Since we are interested in registered voters who voted for either Biden or Trump, we filter the responses that have votereg equal to 1 and cc20_410 equal to 1 or 2. This leaves us with 43,554 responses. Also, because we are interested in the age of the respondents and not their year of birth, we create a new variable called "age" which subtracts the respondent's age from the year of the election, 2020, to get their age. While this may incorrectly represent the age of some people (since we do not have the month and day), for most, it will be an accurate representation of their age given that the election took place on November 4.

Lastly, we then process these responses and decode them to match with the questionnaire responses and obtain [CROSS REFERENCE: TABLE 2].

```{r}
#| include: true
#| warning: false
#| message: false
#| echo: false
# load in libraries.
library(tidyverse)
library(dplyr)
library(rstanarm)
library(modelsummary)

# read and convert values in each column to integer.
ces2020 <-
  read_csv(
    "linear_model_investigation/data/ces2020.parquet",
    col_types =
      cols(
        "votereg" = col_integer(),
        "CC20_410" = col_integer(),
        "gender" = col_integer(),
        "educ" = col_integer(),
        "race" = col_integer(),
        "age" = col_integer()
      )
  )

# formatting data, mapping race ids to races.
ces2020 <-
  ces2020 |>
  filter(votereg == 1,
         CC20_410 %in% c(1, 2)) |>
  mutate(
    voted_for = if_else(CC20_410 == 1, "Biden", "Trump"),
    voted_for = as_factor(voted_for),
    gender = if_else(gender == 1, "Male", "Female"),
    education = case_when(
      educ == 1 ~ "No HS",
      educ == 2 ~ "High school graduate",
      educ == 3 ~ "Some college",
      educ == 4 ~ "2-year",
      educ == 5 ~ "4-year",
      educ == 6 ~ "Post-grad"
    ),
    education = factor(
      education,
      levels = c(
        "No HS",
        "High school graduate",
        "Some college",
        "2-year",
        "4-year",
        "Post-grad"
      )
    ),
    race = case_when(
      race == 1 ~ "White",
      race == 2 ~ "Black",
      race == 3 ~ "Hispanic",
      race == 4 ~ "Asian",
      race == 5 ~ "Native American",
      race == 6 ~ "Middle Eastern",
      race == 7 ~ "Two or more races",
      race == 8 ~ "Other"
    ),
    race = factor(
      race,
      levels = c(
        "White",
        "Black",
        "Hispanic",
        "Asian",
        "Native American",
        "Middle Eastern",
        "Two or more races",
        "Other"
      )
    )
  ) |> select(voted_for, gender, education, race, age)
cat("Table 4. Header for Processed 2020 US Election Survey data.")
# preview data after mutation.
kable(head(ces2020))
```
[CROSS REFERENCE: FIGURE 1] shows that

### 2024-03-17 paper/paper_owais.qmd
test

--- 2024-03-16 paper/paper.qmd
@@
 
-# Simulation
-We first start with simulating the dataset, and recall that we are concerned with each voter's education level, gender, race and age.
-Our chosen sample size is 8000, and the categories for education and race have been named akin to the categories in the dataset.
-Thus, the header of the table obtained for the simulation is shown below:
-```{r}
-#| include: true
-#| warning: false
-#| message: false
-#| echo: false
-# setting seed, prime number :P
-set.seed(71)
-library(knitr)
-library(tibble)
-library(dplyr)
-# sample size 8000
-num_obs <- 8000
-
-# simulating the actual data
-# using 8 categories for races as in the dataset, and ages 0-100.
-us_political_preferences <- tibble(
-  education = sample(0:4, size = num_obs, replace = TRUE),
-  gender = sample(0:1, size = num_obs, replace = TRUE),
-  race = sample(1:8, size = num_obs, replace = TRUE),
-  age = sample(0:100, size = num_obs, replace=TRUE),
-  support_prob = ((education + gender + race + age) / 7),
-) |>
-  mutate(
-    supports_biden = if_else(runif(n = num_obs) < support_prob, "yes", "no"),
-    education = case_when(
-      education == 0 ~ "< High school",
-      education == 1 ~ "High school",
-      education == 2 ~ "Some college",
-      education == 3 ~ "College",
-      education == 4 ~ "Post-grad"
-    ),
-    gender = if_else(gender == 0, "Male", "Female"),
-    race = case_when(
-      race == 1 ~ "White",
-      race == 2 ~ "Black",
-      race == 3 ~ "Hispanic",
-      race == 4 ~ "Asian",
-      race == 5 ~ "Native American",
-      race == 6 ~ "Middle Eastern",
-      race == 7 ~ "Two or more races",
-      race == 8 ~ "Other"
-    )
-  ) |>
-  select(-support_prob, supports_biden, gender, education, race, age)
-
-# preview the simulated data
-cat("Table 1. Header for the political preferences dataset.")
-kable(head(us_political_preferences))
-```
-
 # Data {#sec-data}
@@
 library(dplyr)
+library(knitr)
 library(tidyverse)
@@
 raw_data = read.csv("~/linear_model_investigation/data/ces2020_raw.parquet")
+cat("Table 1. Header for raw election survey data")
 kable(head(raw_data))
@@
 names(raw_data)[names(raw_data) == "birthyr"] <- "age"
-cat("Table 3. Header for 2020 US Election Survey data with the birthyr column converted to ages")
+cat("Table 2. Header for 2020 US Election Survey data with the birthyr column converted to ages")
 kable(head(raw_data))
@@
   ) |> select(voted_for, gender, education, race, age)
-cat("Table 4. Header for Processed 2020 US Election Survey data.")
+cat("Table 3. Header for Processed 2020 US Election Survey data.")
 # preview data after mutation.
@@
 # Model
-Since we wish to predict how likely it is for a respondent to cast a vote for Biden based on their gender, age, education and race, we shall use the rstanarm (citation needed) library to
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+We resort to using a Logistic Regression model since we are interested in predicting the candidate a survey respondent voted for, given the former's age, gender, education and race.
+A logistic regression model is suitable as our response variables (1 for Biden and 0 for Trump) are clearly binary.
 
 ## Model set-up
-We define $y_i$ as the candidate the $ith$ respondent likely voted for, mapping 1 and 0 to Biden and Trump respectively.
-The variables $gender_i$, $age_i$, $race_i$ and $education_i$ are the selected attributes (with the same names, of) of the $ith$ respondent.
-
 \begin{align} 
@@
 
+We define the response variable $y_i$ to be the candidate that the $ith$ respondent voted for.
+Note that the value of $y_i$ is either 0 or 1 indicating whether the candidate in question voted for Trump or Biden respectively, and that $y_i$ is conditioned on $\pi_i$ and follows a Bernoulli distribution with parameter $\pi_i$
+
+The explanatory variables $gender_i$, $age_i$, $race_i$ and $education_i$ are the selected attributes (with the same names, of) of the $ith$ respondent.
+
+We have defined normally distributed priors for the y-intercept $\beta_0$ as well as each of the coefficients of the explanatory variables $\beta_1$, $\beta_2$, $\beta_3$ & $\beta_4$ each with a mean of 0 and standard deviation of 2.5.
+
+Note that the crux of our model, i.e $logit(\pi_i)$ is expressed as a linear combination of our explanatory variables along with the y-intercept $\beta_0$
+
+Thus, the model enables us to estimate whether a candidate voted for Trump or Biden based on their education, gender, race and age.
+The logistic nature of the model ensures that any outputs or response variables obtained are either 0 or 1.
+
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
@@
 ### Model justification
-I actually have no idea what to expect lol
+**Binary Outcome**: logistic regression is specifically designed for binary outcomes, which is suitable for predicting whether a survey respondent voted for Biden (1) or Trump (0).
+Logistic regression models the probability of the response variable falling into a particular category given the predictor variables.
+
+**Interpretability**: Logistic regression provides easily interpretable results.
+The coefficients associated with each predictor variable represent the change in the log-odds of the outcome for a one-unit change in the predictor variable.
+This allows for clear interpretation of the effects of each predictor on the likelihood of voting for a particular candidate.
+
+**Efficiency**: Logistic regression tends to perform well even with relatively small sample sizes compared to more complex models.
+Given that we have a limited set of predictor variables (age, gender, education, and race), logistic regression can efficiently model the relationship between these variables and the voting outcome.
+
+**Regularization Techniques**: If necessary, logistic regression can be extended with regularization techniques such as L1 (Lasso) or L2 (Ridge) regularization to prevent overfitting and improve generalization performance, especially if there are many predictor variables or multicollinearity issues, which can be a useful strategy if overfitting has been observed upon analyzing the results of the model.
+
+**Statistical Inference**: Logistic regression provides inferential statistics such as p-values and confidence intervals for the estimated coefficients, allowing you to assess the significance of each predictor variable in predicting the outcome.
+
+Overall, logistic regression is a robust and interpretable method for modeling binary outcomes like voting preferences, making it a suitable choice for us to indicate what presidential candidate a survey respondent voted for.
 
@@
 
-Our results are summarized in @tbl-modelresults.
-
+Our results are summarized in the table  below:
+&nbsp;
 ```{r}
@@
 #| message: false
-#| eval: true
 #| echo: false
-cat("results")
 political_preferences <-
@@
 # model summary
+cat("Table 4. Whether a respondent is likely to vote for Biden based on their gender, education, race and age (n=8000).")
 modelsummary::modelsummary(
@@
 ```
+
+**Gender**: Test
+
+**Education**: Test
+
+**Race**: Test
+
+**Age**:  Test
 

### 2024-03-17 paper/paper.qmd
added paper

@@
 ---
-title: "Investigating the correlation between a vote for Biden or Trump based on a persons education, gender, race and age."
+title: "Deciphering Key Determinants of Biden Voter Support: Insights from Logistic Regression Analysis in the 2020 US Presidential Election"
 author: 
@@
 
-We are concerned with assessing whether we can predict what candidate a new survey respondent is going to vote for, given their age, education, gender and race.
-Accordingly, we shall be simulating a similar scenario for the 2020 US presidential election below.
+On November 4, 2020, the 59th United States presidential election took place between Joe Biden and Donald Trump, where Joe Biden was elected as the 46th president of the United States.
+Historically, there tend to be common demographics among voters who support the Democratic or Republican party and by extension, the presidential candidates from these parties.
+In addition, voter demographics are strategically considered in presidential campaigns to appeal to certain citizens.
 
-# Simulation
-We first start with simulating the dataset, and recall that we are concerned with each voter's education level, gender, race and age.
-Our chosen sample size is 8000, and the categories for education and race have been named akin to the categories in the dataset.
-Thus, the header of the table obtained for the simulation is shown below:
-```{r}
-#| include: true
-#| warning: false
-#| message: false
-#| echo: false
-# setting seed, prime number :P
-set.seed(71)
-library(knitr)
-library(tibble)
-library(dplyr)
-# sample size 8000
-num_obs <- 8000
+As a result, in this paper the estimand we hope to explore is the likelihood that an American citizen will vote for Joe Biden in the 2020 presidental election given their income, gender, race, and age.
+We will do this by using the 2020 Cooperative Election Study (CES) [CITATION] which is a national online survey that was administered to a representative sample of American citizens pre and post-election, gathering demographic information from 61,000 respondents.
+Using 8000 responses from this survey, we construct a logistic regression model using income, gender, race, and age as predictor variables to see which demographics aligned with support for Joe Biden.
+We find that (**I NEED YOUR MOST RELEVANT FINDINGS HERE**).
+This information is important as it can allow us to understand the campaign strategies in the 2024 United States presidential election toward citizens with specific demographics, especially as Donald Trump and Joe Biden will likely be representatives of their respective parties for the upcoming presidential election.
 
-# simulating the actual data
-# using 8 categories for races as in the dataset, and ages 0-100.
-us_political_preferences <- tibble(
-  education = sample(0:4, size = num_obs, replace = TRUE),
-  gender = sample(0:1, size = num_obs, replace = TRUE),
-  race = sample(1:8, size = num_obs, replace = TRUE),
-  age = sample(0:100, size = num_obs, replace=TRUE),
-  support_prob = ((education + gender + race + age) / 7),
-) |>
-  mutate(
-    supports_biden = if_else(runif(n = num_obs) < support_prob, "yes", "no"),
-    education = case_when(
-      education == 0 ~ "< High school",
-      education == 1 ~ "High school",
-      education == 2 ~ "Some college",
-      education == 3 ~ "College",
-      education == 4 ~ "Post-grad"
-    ),
-    gender = if_else(gender == 0, "Male", "Female"),
-    race = case_when(
-      race == 1 ~ "White",
-      race == 2 ~ "Black",
-      race == 3 ~ "Hispanic",
-      race == 4 ~ "Asian",
-      race == 5 ~ "Native American",
-      race == 6 ~ "Middle Eastern",
-      race == 7 ~ "Two or more races",
-      race == 8 ~ "Other"
-    )
-  ) |>
-  select(-support_prob, supports_biden, gender, education, race, age)
-
-# preview the simulated data
-cat("Table 1. Header for the political preferences dataset.")
-kable(head(us_political_preferences))
-```
+The remainder of this paper is structured as follows: Section 2 discusses the data from the 2020 CES, Section 3 discusses the logistic regression model that was constructed from the 2020 CES data, Section 4 presents the results, and finally Section 5 discusses our findings and some weaknesses.
+(**MOSTLY FROM https://tellingstorieswithdata.com/04-writing_research.html#introduction-1, DUNNO IF IT NEEDS REPHRASING OR GIVEN ADDITIONAL DETAILS LATER**).
 
@@
 
-We will be using the 2020 Cooperative Election Study (CES) (CITATION TODO) as our dataset, which can be previewed below:
+## Data Source
+
+In this paper, we will be using the 2020 Cooperative Election Study (CES) [CITATION], administered by YouGov, is a yearly online survey that gathers information regarding American adults and their political attitudes.
+This study was chosen because its dataset is publicly accessible, includes vote validation, and contains a diverse selection of attributes which make it possible to study how demographics of American adults correlates with preference for presidential candidate.
+
+## Data Measurement
+
+Most, if not all variables in the 2020 CES dataset correspond to a question asked in the CCES online questionnaire.
+Most respondents to this questionnaire are YouGov panelists, but other respondents were recruited from online surveys or other survey providers.
+[CITATION]
+
+However, in order to gain a representative sample of all American adults, not all respondents of the CCES questionnaire end up in the final dataset [CITATION].
+The representative sample was created by using matching, "is a methodology for selection of “representative” samples from non-randomly selected pools of respondent" [CITATION] which is appropriate in this case as the pool of respondents is primarily made of YouGov panelists - which is nonrandom.
+
+Lastly, "individual records were matched to the Catalist database of registered voters in the United States" to validate the votes.
+However, only respondents with a high confidence of being matched in the Catalist database were included.
+Thus, some respondents may have been omitted from the final dataset in this regard.
+
+## Data Features
 
 ```{r}
-#| include: true
 #| warning: false
-#| message: false
 #| echo: false
+#| label: raw_data
+#| tbl-cap: A sample of CES 2020 survey data.
 
@@
 library(tidyverse)
+library(knitr)
 
@@
 ```
-To elaborate further on the dataset starting with the ```votereg``` column, values 1 and 2 signifies whether a person has voted or not, respectively.
-Naturally, for entries with a value of 2 under this column, the corresponding ```CC20_410``` value is NA, and in the other case the values 1 and 2 signify a vote for Biden or Trump respectively.
-The naming of this column corresponds to the question in the survey from which the data for it was collected, which asks "For whom did you vote for President of the United States?" (citation needed), with a total of 6 options, where Biden and Trump correspond to the first two.
 
-Next, there are two values for gender: 1 and 2 corresponding to Male and Female.
-There are also a total of 8 categories under race corresponding to numbers 1 through 8, as well as 6 education categories ranging from "No HS" to "Post-grad".
+We use the R programming language [@citeR] and dataverse [CITATION] to access the CES dataset which initially begins with 61,000 respondents.
+From this dataset, we begin with the following variables as shown in the [FIRST TABLE CROSS REFERENCE].
+The values from these variables from the questionnaire corresponding to each variable can be found in the Appendix along with the value it was numerically coded to in the original dataset [CROSS REFERENCE: TABLE 1].
 
-Since we are interested with ages and not respondent birth years, we may process the dataset as such to reflect this:
+* votereg: This is a binary variable where the values 1 and 2 signify whether a person is registered to vote or not, respectively.
+* cc20_410: This variable measures who the respondent voted for as President of the United States if they voted (with the values 1 and 2 corresponding to a vote for Biden and Trump, respectively) and values 4 through 7 that distinguish non-voters
+* gender: This is a binary variable which measures if the respondent is male (1) or female (2)
+* educ: This is a nominal variable which measures the level of education the respondent has ranging from 1 - "Did not graduate from high school" to 6 - "Postgraduate degree"
+* race: This is a nominal variable which numerically codes races from the questionnaire to numbers in the dataset
+* birthyr: This variable measures the year of birth of a respondent.
+
+Since we are interested in registered voters who voted for either Biden or Trump, we filter the responses that have votereg equal to 1 and cc20_410 equal to 1 or 2.
+This leaves us with 43,554 responses.
+Also, because we are interested in the age of the respondents and not their year of birth, we create a new variable called "age" which subtracts the respondent's age from the year of the election, 2020, to get their age.
+While this may incorrectly represent the age of some people (since we do not have the month and day), for most, it will be an accurate representation of their age given that the election took place on November 4.
+
+Lastly, we then process these responses and decode them to match with the questionnaire responses and obtain [CROSS REFERENCE: TABLE 2].
 
 ```{r}
-#| include: true
 #| warning: false
-#| message: false
 #| echo: false
-
-raw_data$birthyr <- as.numeric(raw_data$birthyr)
-current_year <- 2020
-
-# Subtract current year from birth year column to get age
-raw_data <- raw_data %>% mutate(birthyr = current_year - birthyr)
-# rename to age
-names(raw_data)[names(raw_data) == "birthyr"] <- "age"
-cat("Table 3. Header for 2020 US Election Survey data with the birthyr column converted to ages")
-kable(head(raw_data))
-
-# save data back into the csv.
-write.csv(raw_data,"~/linear_model_investigation/data/ces2020.parquet")
-```
-We shall now filter for non NA values within the ```CC20_410``` column, as we are only concerned with respondents that cast a vote, which also means we filter for entries with a ```votereg``` value of 1.
-Additionally, we shall map each education level and race value to its corresponding value from the survey, and we finally obtain a table as follows:
-
-```{r}
-#| include: true
-#| warning: false
-#| message: false
-#| echo: false
+#| label: final_data
+#| tbl-cap: Header for Processed 2020 CES survey data
 # load in libraries.
@@
   ) |> select(voted_for, gender, education, race, age)
-cat("Table 4. Header for Processed 2020 US Election Survey data.")
 # preview data after mutation.
@@
 ```
-TODO: complete bar chart section from the worked example.
+[CROSS REFERENCE: FIGURE 1] shows that
 
 # Model
-Since we wish to predict how likely it is for a respondent to cast a vote for Biden based on their gender, age, education and race, we shall use the rstanarm (citation needed) library to
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+We resort to using a Logistic Regression model since we are interested in predicting the candidate a survey respondent voted for, given the former's age, gender, education and race.
+A logistic regression model is suitable as our response variables (1 for Biden and 0 for Trump) are clearly binary.
 
 ## Model set-up
-We define $y_i$ as the candidate the $ith$ respondent likely voted for, mapping 1 and 0 to Biden and Trump respectively.
-The variables $gender_i$, $age_i$, $race_i$ and $education_i$ are the selected attributes (with the same names, of) of the $ith$ respondent.
-
 \begin{align} 
@@
 
-Our results are summarized in @tbl-modelresults.
-
+Our results are summarized in the table  below:
+&nbsp;
 ```{r}
-#| include: true
 #| warning: false
-#| message: false
-#| eval: true
 #| echo: false
-cat("results")
+#| label: model_summary
+#| tbl-cap: Whether a respondent is likely to vote for Biden based on their gender, education, race and age (n=8000).
 political_preferences <-

### 2024-03-17 paper/paper.qmd
done reuslts

@@
 ```{r}
-#| warning: false
 #| echo: false
@@
 ```
+As a note, the standard error for any value discussed below shall be abbreviated as sd-err.
 
-**Gender**: Test
+**Gender**: The coefficient for Males is -0.45 (0.051 sd-err) which means Males are less likely to cast a vote for Biden compared to Females, or that Males would choose Trump in an election, given that Trump and Biden were the only candidates to choose from.
 
-**Education**: Test
+**Education**: We see that with an increase in literacy level, the likelihood that a candidate will vote for Biden increases.
+Observe that as we move upwards in the following list of categories: High school graduate, some college, 2-year, 4-year and post-grad, we see that the corresponding coefficients are -0.267 (sd-err 0.198), 0.054 (sd-err 0.195), 0.095 (sd-err 0.203), 0.445 (sd-err 0.196) and 0.906 (sd-err 0.199) respectively, meaning that the more educated a candidate is, the more likely they are to vote for Biden.
 
-**Race**: Test
+**Race**: Amongst the positive coefficients, observe that voters identifying on the survey as Black demonstrate the highest support for Biden amongst all other racial groups of respondents with a coefficient of 2.590 (sd-err 0.155), whereas Hispanic and Asian respondents bear coefficients 0.572 (sd-err 0.905) and 0.635 (sd-err 0.162) respectively.
+Coming in last place, we have Middle Eastern respondents with the smallest positive coefficient compared to the other racial groups, with a value of 0.356 (sd-err 0.181).
+Finally, we have respondents identifying with "Two or more races" that are actually LESS likely to support Biden, with a value of -0.294 (sd-err 0.172).
+Finally, we observe that respondents identifying as "Other" demonstrate a coefficient of 1.301, although with a comparatively higher standard error of 1.277.
 
-**Age**:  Test
+**Age**: We see a coefficient value of -0.018 (sd-err 0.002) showing a weak negative correlation between the age of a respondent and whether they voted for Biden.
+All in all, it is bold to claim that there exists a definite negative correlation between a voter's age and whether they voted for Biden.
 

### 2024-03-17 paper/paper.qmd
test
added stats table

@@
 date-format: long
-abstract: "In this paper, we discuss the correlation between various statistics of the respondents to the CES2020 election survey, and whether they voted for Trump or Biden.
-The statistics are: age, education, gender and race."
+abstract: "On November 4, 2020, the 59th United States presidental election took place where Joe Biden became elected as 46th president of the United States.
+Using responses from the CES 2020 election survey (n = 8000), we study the effect that gender, race, age, and education have on whether an American voted for Joe Biden using a logistic regression model.
+We find that men are less likely to vote for Biden than women, respondents with higher education level are more likely to vote for Biden, and that Black respondents are most likely to vote for Biden from all other racial groups.
+We also find a weak negative correlation between a respondents age and whether they vote for Biden.
+As the upcoming United States presidental election is anticipated to be between Donald J.
+Trump and Joe Biden, these results have implications for how both candidates will use voter demographics to guide their campaigns."
 format: pdf
@@
 ```{r}
+#| warning: false 
 #| echo: false
@@
 [CROSS REFERENCE: FIGURE 1] shows that
+```{r}
+#| warning: false
+#| echo: false
+#| label: dist
+#| tbl-cap: The distribution of presidential preferences, by race and education.
+ces2020 |>
+  ggplot(aes(x = education, fill = voted_for)) +
+  stat_count(position = "dodge") +
+  facet_wrap(facets = vars(race)) +
+  theme_minimal() +
+  labs(
+    x = "Highest education",
+    y = "Number of respondents",
+    fill = "Voted for"
+  ) +
+  coord_flip() +
+  scale_fill_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+```
 

### 2024-03-17 paper/paper.qmd
Paper Updates

@@
   - Owais Zahid
-thanks: "Code and data are available at: https://github.com/FFFiend/linear_model_investigation."
+thanks: "Code and data are available at: https://github.com/FFFiend/us_2020_election_analysis."
 date: today
@@
 Using 8000 responses from this survey, we construct a logistic regression model using income, gender, race, and age as predictor variables to see which demographics aligned with support for Joe Biden.
-We find that (**I NEED YOUR MOST RELEVANT FINDINGS HERE**).
+We find that men are less likely to vote for Biden than women, respondents with higher education level are more likely to vote for Biden, and that Black respondents are most likely to vote for Biden from all other racial groups.
+We also find a weak negative correlation between a respondents age and whether they vote for Biden.
 This information is important as it can allow us to understand the campaign strategies in the 2024 United States presidential election toward citizens with specific demographics, especially as Donald Trump and Joe Biden will likely be representatives of their respective parties for the upcoming presidential election.
 
-The remainder of this paper is structured as follows: Section 2 discusses the data from the 2020 CES, Section 3 discusses the logistic regression model that was constructed from the 2020 CES data, Section 4 presents the results, and finally Section 5 discusses our findings and some weaknesses.
-(**MOSTLY FROM https://tellingstorieswithdata.com/04-writing_research.html#introduction-1, DUNNO IF IT NEEDS REPHRASING OR GIVEN ADDITIONAL DETAILS LATER**).
+The remainder of this paper is structured as follows: @sec-data discusses the data from the 2020 CES, @sec-model discusses the logistic regression model that was constructed from the 2020 CES data, and @sec-results presents the results.
 
@@
 
-# Model
+# Model {#sec-model}
 We resort to using a Logistic Regression model since we are interested in predicting the candidate a survey respondent voted for, given the former's age, gender, education and race.
@@
 
-# Results
+# Results {#sec-results}
 
@@
 
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-## Diagnostics
-
-\newpage
-
 # References

### 2024-03-17 paper/paper.qmd
usage + paper

@@
 
+We shall be using the the following libraries: dplyr [@citedplyr], tidyverse [@citetidyverse], knitr [@citeknitr], stanarm [@citestanarm] and modelsummary [@citemodelsummary].
+
 ## Data Measurement
@@
 
-We use the R programming language [@citeR] and dataverse [CITATION] to access the CES dataset which initially begins with 61,000 respondents.
+We use the R programming language [@citeR] and dataverse [@citedataverse] to access the CES dataset which initially begins with 61,000 respondents.
 From this dataset, we begin with the following variables as shown in the [FIRST TABLE CROSS REFERENCE].
